\begin{cabstract}
在临时布防、应急处置、移动巡检和现场安保等场景中，布控球通常需要在比较复杂的环境下完成目标发现、持续跟踪和云台自适应控制。传统布控球更多依靠人工遥控或固定巡航策略，当目标移动较快、多人同时进入画面、目标被短时遮挡，或者设备采用倒挂安装方式时，系统容易出现响应不及时、跟踪过程有抖动、目标切换不够稳定等问题。如果把视频流持续上传到云端进行分析，又会受到网络带宽、通信时延、隐私保护和现场部署条件的影响。基于这些问题，本文研究并实现了一种基于海康威视设备、YOLO 目标检测和自适应 PTZ 控制的嵌入式云台跟踪系统，用于提高布控球在边缘端的智能化水平。

系统实现过程中，本文完成了海康威视 HCNetSDK 与 PlayCtrl 播放库的 Python 封装和调用，搭建了从设备登录、实时预览、码流解码到帧缓存读取的视频采集链路；在目标感知部分，系统选用轻量级 YOLOv8n 模型进行人体目标检测，并结合置信度过滤、坐标提取和人体外观特征构建，实现对布控球画面的实时目标分析。在控制层面，系统根据人体目标中心与画面中心之间的偏差设计 PTZ 控制策略，把偏差阈值、比例脉冲、控制冷却和指数平滑机制结合起来，使云台在保持响应速度的同时尽量减少频繁抖动。针对多人同时出现的情况，本文设计了轻量级人体 ReID 记忆模块，为检测到的人体目标分配相对稳定的编号，支持自动选择最大人体目标，也支持手动指定目标 ID，并在目标短暂离开画面后继续等待其重新入镜。系统还对倒挂安装、人工接管、检测结果反馈和异常退出等实际使用问题进行了适配，并构建了人机交互模块，用来完成视频预览、目标状态呈现、模式切换、人工接管和检测结果反馈等功能。

初步测试结果显示，系统能够较稳定地获取布控球实时视频流，并完成人体检测、目标编号、自动/指定目标跟踪和云台方向控制。YOLOv8n 在当前配置下能够满足实时检测需求，ReID 记忆模块增强了多人场景下的目标连续性，平滑控制策略也能改善云台跟踪时的抖动现象。本文把视觉检测、目标记忆、云台控制和人机交互集成到真实布控球设备中，为低成本布控球智能跟踪设备提供了一种可实现的方案，也为模型量化、场景增强、嵌入式平台迁移和云边协同等工程拓展提供了基础。
\end{cabstract}

\begin{eabstract}
In temporary surveillance, emergency response, mobile inspection, and on-site security scenarios, a ball camera is often required to detect targets, maintain tracking, and control its pan-tilt unit adaptively under complex conditions. Conventional ball camera systems mainly rely on manual operation or fixed patrol routes. When the target moves quickly, multiple people appear at the same time, short-term occlusion occurs, or the device is installed upside down, the system may suffer from delayed response, tracking jitter, and unstable target switching. If the video stream is continuously uploaded to the cloud for analysis, the system will also be affected by network bandwidth, communication latency, privacy concerns, and deployment conditions. To address these issues, this thesis studies and implements an embedded adaptive PTZ tracking system based on Hikvision devices, YOLO object detection, and visual feedback control.

During system implementation, Hikvision HCNetSDK and the PlayCtrl library are wrapped and invoked in Python, so that a video acquisition pipeline from device login, real-time preview, stream decoding, to frame buffering can be built. A lightweight YOLOv8n model is used for human detection, and real-time target analysis is carried out through confidence filtering, coordinate extraction, and lightweight appearance feature construction. At the control layer, the system calculates the deviation between the human target center and the image center, and then generates PTZ commands by combining dead-zone thresholds, proportional pulse duration, command cooldown, and exponential smoothing. These mechanisms allow the pan-tilt unit to respond in time while reducing unnecessary oscillation. For multi-person scenes, a lightweight body ReID memory module is designed to assign relatively stable IDs to detected people. The system supports automatic selection of the largest human target and manual tracking of a specified person ID, and it can wait for the selected target when the target temporarily leaves the frame. Engineering adaptations are also made for inverted installation, manual intervention, detection feedback, and abnormal exit handling. In addition, a human-computer interaction module is developed to support video preview, target-state presentation, mode switching, manual takeover, and detection feedback.

Preliminary tests show that the system can obtain real-time video streams from the ball camera with stable performance, and can complete human detection, identity assignment, automatic or specified target tracking, and PTZ directional control. YOLOv8n meets the real-time detection requirement under the current configuration, the ReID memory module enhances target continuity in multi-person scenes, and the smoothed control strategy helps reduce pan-tilt jitter. By integrating visual detection, target memory, PTZ control, and human-computer interaction on a real ball camera, this work provides a practical implementation path for low-cost intelligent tracking devices and lays an engineering foundation for model quantization, scene enhancement, embedded platform migration, and edge-cloud collaborative optimization.
\end{eabstract}
